### 2018-07-11 draft.Rmd
add last version

coord <- read_tsv("data/coordinates.tsv")

@@

coord %>% 
  filter(Lat > 42.6,
         Lat < 42.9,
         Lon > 46.1,
         Lon < 46.5) %>% 
  mutate(label = if_else(is.na(label), "", label)) ->
  coord

@@

```{r, fig.width=6, fig.height=4.5, cashe = TRUE, fig.cap= "Figure 2. The Andi villages (labeled) and their neighbours"}

coord$Language <- factor(coord$Language, levels = c("Chechen", "Avar", "Andi", "Botlikh", "Godoberi", "Karata"))

map.feature(coord$Language,
            features = coord$Language,
            latitude = coord$Lat,
            longitude = coord$Lon,
            width = 6,
            label.emphasize = list(24, "red"),
            color = c("darkgrey", "deepskyblue", "darkseagreen", "aquamarine4", "aquamarine2", "lightblue1"),
            label = coord$label,
            label.hide = FALSE,
            label.fsize = 14,
            label.font = 'Brill',
            zoom.level = 11,
            tile = c("Esri.WorldGrayCanvas"))
```

@@

As mentioned earlier, the attribution of the masculine and feminine noun classes is very transparent. The animate class consists of animals, both large and small (the only exception are insects, which in some cases are grouped with inanimates). The motivation for assigning any of the inanimate classes remains a mystery. Not even all insects belong to the same inanimate class: in Rikvani, spiders belong to the animate class, while butterflies belong to the third inanimate class. In addition, *aqrab* 'scorpion', which is a loan from Avar, also belongs to the animate class, while the native word *hortʃ'oritʃ'in* 'scorpion, belongs to the third inanimate class. A possible explanation for *aqrab* belonging to inanimate 1 is analogy with Avar - in Avar there is only one neuter class, which comprises animals and objects, and corresponds to the class marker *b-*, parallel to animate and inanimate 1 in the Andi dialects.

@@

```{r, dpi = 500, cashe = TRUE, fig.cap= "Figure 3. Dendrogram showing the clusterization of speakers."}
df %>% 
  mutate(class = if_else(class == "b", 1, 0)) %>% 
  select(-sex, -age_2017, -group_age) %>% 
  spread(s_id, class) %>% 
  select(`1`:`16`) %>% 
  t() ->
  df_cluster

@@

# Гарик, тут что-то не то.. бабочка у нас комната

@@

* the word *haq'u* 'butterfly' was attributed inanimate 1 class by older speakers four times, and by younger speakers five times, so its variability rate will be different in different groups: $$\frac{inanimate_1}{inanimate_1+inanimate_2} = \frac{4}{4+4} = 0.5$$
$$\frac{inanimate_1}{inanimate_1+inanimate_2} = \frac{5}{5+3} = 0.625$$

@@

```{r, dpi = 500, cashe = TRUE, fig.cap= "Figure 4. Destribution of the most variable words by different age group.", fig.width=7}
df %>%  
  group_by(w_id, translation_en, stimulus_source, group_age) %>% 
  count(class) %>% 
  spread(class, n, fill = 0) %>% 
  mutate(rate_b = b/(b+r),
         repel = if_else(rate_b > 0 & rate_b < 1, TRUE, FALSE)) %>%
  filter(repel == TRUE) %>%
  select(-b, -r) %>%
  spread(group_age, rate_b, fill = 0)  %>% 
  ggplot(aes(younger, older))+
  geom_abline(slope = 1, intercept = 0, linetype  = 2)+
  geom_point()+
  geom_point(data = data.frame(younger = 0:1, older = 0:1),
             aes(younger, older))+
  ggrepel::geom_text_repel(aes(label = translation_en, color = stimulus_source), show.legend = FALSE, seed = 42, family = "Times New Roman")+
  scale_x_continuous(limits = c(0, 1))+
  scale_y_continuous(limits = c(0, 1))+
  labs(x = expression(paste("rate of ", italic("1 inanimated"), "class (by ", underline(bold(younger))," spekers)")),
       y = expression(paste("rate of ", italic("1 inanimated"), "class (by ", underline(bold(older)), " spekers)")))
```

>>>>

#coord <- read_tsv("data/coordinates.tsv")

@@

# coord %>%
#   filter(Lat > 42.6,
#          Lat < 42.9,
#          Lon > 46.1,
#          Lon < 46.5) %>%
#   mutate(Language = factor(Language, levels = c("Chechen", "Avar", "Andi", "Botlikh", "Godoberi", "Karata"))) ->
#   coord
# 
#write_tsv(coord, "data/coordinates_2.tsv", na = "")
coord <- read_tsv("data/coordinates_2.tsv")
coord %>% 
  mutate(label = if_else(is.na(label), "", label)) ->
  coord

@@

```{r, fig.width=6, fig.height=4.5, cashe = TRUE, fig.cap= "Figure 2. The Andi villages (labeled) and their neighbours"}
map.feature(coord$Language,
            features = coord$Language,
            latitude = coord$Lat,
            longitude = coord$Lon,
            width = 6,
            label.emphasize = list(37, "red"),
#            color = c("darkgrey", "deepskyblue", "darkseagreen", "aquamarine4", "aquamarine2", "lightblue1"),
            label = coord$label,
            label.hide = FALSE,
            label.fsize = 14,
            label.font = 'Brill',
            zoom.level = 11,
            tile = c("Esri.WorldGrayCanvas"))
```

@@

As mentioned earlier, the attribution of the masculine and feminine noun classes is very transparent. The animate class consists of animals, both large and small (the only exception are insects, which in some cases are grouped with inanimates). The motivation for assigning any of the inanimate classes remains a mystery. Not even all insects belong to the same inanimate class: in Rikvani, spiders belong to the animate class, while butterflies belong to the third inanimate class. In addition, *aqrab* 'scorpion', which is a loan from Avar, also belongs to the animate class, while the native word *hortʃ'oritʃ'in* 'scorpion, belongs to the third inanimate class. A possible explanation for *aqrab* belonging to inanimate 1 is analogy with Avar --- in Avar there is only one neuter class, which comprises animals and objects, and corresponds to the class marker *b-*, parallel to animate and inanimate 1 in the Andi dialects. In Figure 3. the number of classes in Andi and neighbouring villages is presented.

```{r, fig.width=6, fig.height=4.5, cashe = TRUE, fig.cap= "Figure 3. Number of classes in Andi and their neighbours."}

map.feature(coord$Language,
            features = paste(coord$class, "classes"),
            latitude = coord$Lat,
            longitude = coord$Lon,
            width = 6,
            label.emphasize = list(37, "red"),
            label = coord$label,
            label.hide = FALSE,
            label.fsize = 14,
            label.font = 'Brill',
            zoom.level = 11,
            tile = c("Esri.WorldGrayCanvas"))
```

@@

```{r, dpi = 500, cashe = TRUE, fig.cap= "Figure 4. Dendrogram showing the clusterization of speakers."}
df %>% 
  mutate(class = if_else(class == "b", 1, 0)) %>% 
  select(-sex, -age_2017, -group_age) %>% 
  spread(s_id, class) %>% 
  select(`1`:`16`) %>% 
  t() ->
  df_cluster

@@

@@

* the word *odoruk'a* 'butterfly' was attributed inanimate 1 class by older speakers four times, and by younger speakers five times, so its variability rate will be different in different groups: $$\frac{inanimate_1}{inanimate_1+inanimate_2} = \frac{4}{4+4} = 0.5$$
$$\frac{inanimate_1}{inanimate_1+inanimate_2} = \frac{5}{5+3} = 0.625$$

@@

```{r, dpi = 500, cashe = TRUE, fig.cap= "Figure 5. Distribution of the most variable words by different age group (native lexemes are in blue).", fig.width=7}
df %>%  
  group_by(w_id, translation_en, stimulus_source, group_age) %>% 
  count(class) %>% 
  spread(class, n, fill = 0) %>% 
  mutate(rate_b = b/(b+r),
         repel = if_else(rate_b > 0 & rate_b < 1, TRUE, FALSE)) %>%
  filter(repel == TRUE) %>%
  select(-b, -r) %>%
  spread(group_age, rate_b, fill = 0)  %>% 
  ggplot(aes(younger, older))+
  geom_abline(slope = 1, intercept = 0, linetype  = 2)+
  geom_point()+
  geom_point(data = data.frame(younger = 0:1, older = 0:1),
             aes(younger, older))+
  ggrepel::geom_text_repel(aes(label = translation_en, color = stimulus_source), show.legend = FALSE, seed = 42, family = "Times New Roman")+
  scale_x_continuous(limits = c(0, 1))+
  scale_y_continuous(limits = c(0, 1))+
  labs(x = expression(paste("rate of ", italic("1 inanimated"), "class (by ", underline(bold(younger))," spekers)")),
       y = expression(paste("rate of ", italic("1 inanimated"), "class (by ", underline(bold(older)), " spekers)")))
```
